---
title: "In-class Exercise 5"
author: "Ee Fook Ming"
date: "February 15, 2025"
date-modified: last-modified
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
format: 
  html:
    number-sections: true
#    number-offset: [5, 0]  # Offset heading levels by starting at 6
output: 
  html_document: 
    toc: true
    tabset: true
---


# In-class Exercise 05

## Getting Started

```{r}

# first check if the package is installed, if not installation kicks off. Then it loads the package
pacman::p_load(tidyverse, readxl, SmartEDA, easystats, gtsummary, ggstatsplot)

```


## Importing Data

```{r}

# Read the excel file and pick up the data in worksheet called "data"
car_resale <- read_xls("data/ToyotaCorolla.xls", "data")


```


## Data Overview

```{r}

car_resale %>%
  ExpData(type = 1)

summary(car_resale)

# glimpse(car_resale)

```

```{r}
list(car_resale)
```

```{r}
summary(car_resale)
```


## Using SmartEDA - Productivity Tool for EDA Analysis

### Type = 1

```{r}

# Using SmartEDA - productive tool for EDA analysis

# summary1 <- car_resale %>%
car_resale %>%
  ExpData(type = 1)  # Explore data with type 1. There is type 2.

```


### Type = 2

```{r}

# Using SmartEDA - productive tool for EDA analysis

car_resale %>%
  ExpData(type = 2)  # Explore data with type 1. There is type 2.

```


## Data Loading and Preprocessing for Car Resale Analysis

```{r}

cols <- c("Mfg_Month", "HP_Bin", "CC_bin", "Doors", "Gears",
          "Cylinders", "Fuel_Type", "Color",
          "Met_Color", "Automatic", "Mfr_Guarantee", "BOVAG_Guarantee",
          "ABS", "Airbag_1",
          "Airbag_2", "Airco", "Automatic_airco", "Boardcomputer",
          "CD_Player",
          "Central_Lock", "Powered_Windows", "Power_Steering", "Radio",
          "Mistlamps",
          "Sport_Model", "Backseat_Divider", "Metallic_Rim",
          "Radio_cassette", "Tow_Bar")

car_resale <- read_xls("data/ToyotaCorolla.xls", sheet = "data") %>%
  mutate(Id = as.character(Id)) %>%
  mutate(across(all_of(cols), as.factor))


```


## Exploratory Numeric Visualization of Car Resale Data


```{r}
car_resale %>%
  ExpNumViz(target=NULL,
            nlim=10,
            Page=c(2,2))

```
## Exploratory Numeric Visualization of Car Resale Data with Price as Target


```{r}
car_resale %>%
  ExpNumViz(target="Price",
            nlim=10,
            Page=c(2,2)) #, theme="default")
```

## Bar Plots for All Categorical Variables

```{r}
#| fig-width: 12
#| fig-height: 10

car_resale %>%
  ExpCatViz(target=NULL,
            col = "sky blue",
            clim=10,
            margin=2,
            Page=c(4,4),
            sample=16)

```



## easystats - parameters, performance and see

### Using parameters to output of the model's parameters - check multi-colinearity in the data

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + Weight + Guarantee_Period, data = car_resale)  # lm = linear regression model

model
```

```{r}

check_collinearity(model)

```


```{r}
check_c <-check_collinearity(model)
plot(check_c)
```


### Check Normality
```{r}

model1 <- lm(Price ~ Age_08_04 + KM + Weight + Guarantee_Period, data = car_resale)
check_normality(model1)

```

```{r}
check_c1 <-check_normality(model1)
plot(check_c1)
```


### Check Heteroscedasticity

```{r}

check_c2 <-check_heteroscedasticity(model1)
plot(check_c2)


```


### Check All

```{r}
#| fig-width: 12
#| fig-height: 8


check_c_all <- check_model(model1)
plot(check_c_all)

```


### Regression Model Summary with Intercept


```{r}

# summary(model1)
tbl_regression(model1,intercept = TRUE) # summarize regression model


```


## Extracting Model Parameters and the Plot

```{r}
p_model <- parameters((model1))
```


```{r}
# plot(parameters((model1)))
plot(p_model)

```


### Coefficient Plot for Model1 using ggcoefstats


```{r}

ggcoefstats(model1, 
            output = "plot")

```







