---
title: "Hands-on Exercise 00: working with tidyverse"
author: "Ee Fook Ming"
date-modified: "last-modified"
execute:
  echo: true
  eval: true
  warning: false
  freeze: true
---

## Getting started

Load tidyverse onto r environment by using the code chunk below.

```{r}

pacman::p_load(tidyverse)

```

## Importing data

Code chunk below uses read_csv() or readr to import REALIS2019.csv into environment as a tibble data.frame

read.csv is the native R read. It should not be used as it convert space to dot for the field name
```{r}
# #| eval: FALSE
realis_csv <- read.csv("data/REALIS2019.csv")

```

```{r}
realis2019 <- read_csv("data/REALIS2019.csv")

```

```{r}
# #| eval: FALSE
popdata_fat <- read_csv("data/PopData2019_fat.csv")


```

## Pivoting Data

```{r}
# #| eval: FALSE
popdata_long <- popdata_fat %>%
  pivot_longer(c(3:21),
        names_to = "Age Group",
        values_to = "Population")

```

```{r}
# #| eval: FALSE
write_rds(popdata_long,"data/rds/popdata_long.rds")

```

```{r}
# #| echo: FALSE
popdata_long <- read_rds("data/rds/popdata_long.rds")

```

## Working with dplyr

```{r}
realis2019_selected <- realis2019 %>% 
  select('Project Name',
  'Transacted Price ($)',
  'Type of Sale',
  'Unit Price ($ psm)',
  'Property Type')
realis2019_selected
```

### Working with filter() of dplyr

```{r}

realis2019_filtered <- realis2019_selected %>%
  filter('Property Type' == "Condominium" |
           'Property Type' == "Apartment") %>%
  filter('Type of Sale' == "New Sale") %>%
  filter('Unit Price ($ psm)' <= 13000)

```

### Putting it all together

```{r}

realis2019_end <- realis2019 %>% 
  select('Project Name',
  'Transacted Price ($)',
  'Type of Sale',
  'Unit Price ($ psm)',
  'Property Type') %>%
  filter('Property Type' == "Condominium" |
           'Property Type' == "Apartment") %>%
  filter('Type of Sale' == "New Sale") %>%
  filter('Unit Price ($ psm)' <= 13000)

  
```
